---
title: "Code Draft Rmd"
author: "Lyttia Cancinos-McManus"
date: "July 13, 2018"
output: html_document
---

# Capstone Project Data Wrangling  

_I will be using an R interface to weather underground's API. In order to use this library you must obtain an API key. In the past, a free key was available. The free key limits requests to 10 per minute. If you are grabbing weather for a large date range using history_range then by default limit = 10 will limit the calls to a maximum of 10 per minute. This package has functions that follow the online api._

**0) Install & Load packages**

```{r}
#install.packages("tidyverse")
library(tidyverse)
#install.packages("devtools")
library(devtools)
#devtools::install_github("ALShum/rwunderground", force = TRUE)
library(rwunderground)
```

**0a) Set Variables**

```{r}
rwunderground::set_api_key("337b40a1234484ca")
WUNDERGROUNDID = '337b40a1234484ca'
```

**1) Read and View original data (csv files)**

```{r}
PropVars_original <- read_csv("RDC_InventoryCoreMetrics_Zip_Hist.csv")
View(PropVars_original)

Crime_df_original <- read_csv("Phoenix Crime Data 7.2.18.csv")
View(Crime_df_original)

Crime_df_refine <- separate(Crime_df_original, 2, c("Start Date", "Start Time"), sep = " ")
Crime_df_refine <- separate(Crime_df_refine, 4, c("End Date", "End Time"), sep = " ")

View(Crime_df_refine)
```

**1a) Call Wundergound API to create the PHX weather history df to be analyzed**

```{r}
ZIPCODE = (as.character(unique(Crime_df_refine$ZIP)))

location_zips = sapply(ZIPCODE, set_location)

getdata <- function(location_zip,date_start = "20151101", date_end = "20151102")

{ df  = history_range(location_zip,date_start,date_end)
  return (df) }

df_combined = do.call(rbind, lapply(location_zips, getdata))

View(df_combined)
```

*NOTE: The dates used above are TEST DATES to limit calls while experimenting.*

**1b) Clean Variables/Column Names/Observations** 

```{r}
PropVars_refine <- PropVars_original[ which (PropVars_original$ZipName == "Phoenix, AZ" & PropVars_original$Month >= 2015-11-01), ]
```

*NOTE: The above code does not successfully remove observations (PropVars_original$Month aka:dates) before 2015-11-01. Need to edit date format. Maybe convert column to different(unknown) date format before trying code above.* 

```{r}
PropVars_refine <- PropVars_refine %>% select('ZipName', 'Month', 'ZipCode', 'Median Listing Price', 'Avg Listing Price', 'Total Listing Count')
```

*Find the distribution of listing prices to determine whether to use median or mean in PropVars_refine df.* 

```{r}
View(PropVars_refine)
```
