---
title: "Data Wrangling Script (in progress)"
author: "Lyttia Cancinos-McManus"
date: "July 13, 2018"
output: html_document
---

# Capstone Project Data Wrangling  

_I will be using an R interface to weather underground's API. In order to use this library you must obtain an API key. Prior to ____, a free key was available. The free key limits requests to 10 per minute. If you are grabbing weather for a large date range using history_range then by default limit = 10 will limit the calls to a maximum of 10 per minute. This package has functions that follow the online api._

**0) Install & Load packages**

```{r}
#install.packages("tidyverse")
#install.packages("devtools")
library(tidyverse)
library(devtools)
#devtools::install_github("ALShum/rwunderground", force = TRUE)
library(rwunderground)
```

**0a) Set API Key**

```{r}
rwunderground::set_api_key("337b40a1234484ca")
WUNDERGROUNDID = '337b40a1234484ca'
```

**1) Read and examine original data (csv files)**

_Property data original csv is too large to upload to github. Access the original csv file from [Realtor.com](https://www.realtor.com/research/data)._

```{r}
propvals_original <- read_csv("RDC_InventoryCoreMetrics_Zip_Hist.csv")
propvals <- tbl_df(propvals_original)
propvals
str(propvals)
head(propvals)
data.frame(head(propvals))

crimes_original <- read_csv("Phoenix Crime Data 7.2.18.csv")
crimes <- tbl_df(crimes_original)
crimes
str(crimes)
head(crimes)
data.frame(head(crimes))
```

**1a) Call rwundergound API Package to create PHX weather history df**

```{r}
zip_code = (as.character(unique(crimes$ZIP)))
location_zips = sapply(zip_code, set_location)
getdata <- function(location_zip, date_start = "20151101", 
                    date_end = "20151102"){
                    df  = history_range(location_zip,date_start,date_end) 
                    return (df)
                    }
weather_df_combined = do.call(rbind, lapply(location_zips, getdata))
View(weather_df_combined)
weather_df_combined <- tbl_df(weather_df_combined)
```

*NOTE: The dates used above are TEST DATES to limit calls while experimenting.*

**2) Clean Variables/Column Names/Observations** 

```{r}
propvals <- propvals[ which(propvals$ZipName == "Phoenix, AZ" & 
                            propvals$Month >= "2015-10-01"), ]  
propvals <- propvals %>% 
  select('Month', 'ZipCode', 'Median Listing Price', 'Total Listing Count') 

View(propvals)
```

**2a) Separate date and time in crimes df**

```{r}
crimes <- separate(crimes, 2, c("start_date", "start_time"), sep = " ")
crimes <- separate(crimes, 4, c("end_date", "end_time"), sep = " ")
View(crimes)
```
**Coming Soon:
 1. Propvals df: Rename- month, year, zip_code, median_value, total_listing.
 2. Crimes df: Remove- 100 Block Addr Crime, Separate- dates col to month, year, day.
 3. Possibly create dummy variables for premise type and crime type (ucr crime category)**
