---
title: "Data Wrangling Script (in progress)"
author: "Lyttia Cancinos-McManus"
date: "July 13, 2018"
output: html_document
---

# Capstone Project Data Wrangling  

_I will be using an R interface to weather underground's API. In order to use this library you must obtain an API key. Prior to ____, a free key was available. The free key limits requests to 10 per minute. If you are grabbing weather for a large date range using history_range then by default limit = 10 will limit the calls to a maximum of 10 per minute. This package has functions that follow the online api._

**0) Install & Load packages**

```{r}
install.packages("tidyverse", dependencies = TRUE)
install.packages("devtools")
install.packages("stringi", dependencies = TRUE)
library(tidyverse)
library(devtools)
devtools::install_github("ALShum/rwunderground", force = TRUE)
library(rwunderground)
```

**0a) Set API Key**

```{r}
rwunderground::set_api_key("337b40a1234484ca")
WUNDERGROUNDID = '337b40a1234484ca'
```

**1) Read and examine original data (csv files)**

_Property data original csv is too large to upload to github. Access the original csv file from [Realtor.com](https://www.realtor.com/research/data)._

```{r}
property_original <- read_csv("RDC_InventoryCoreMetrics_Zip_Hist2.csv")
property <- tbl_df(property_original)
str(property)
head(property)
data.frame(head(property))

crimes_original <- read_csv("Phoenix Crime Data 7.2.18.csv")
crimes <- tbl_df(crimes_original)
str(crimes)
head(crimes)
data.frame(head(crimes))
```

**1a) Call rwundergound API Package to create PHX weather history df**

```{r}
zip_code = (as.character(unique(crimes$ZIP)))
location_zips = sapply(zip_code, set_location)
getdata <- function(location_zip, date_start = "20151101", 
                    date_end = "20151102"){
                    df  = history_range(location_zip,date_start,date_end) 
                    return (df)
                    }
weather_df_combined = do.call(rbind, lapply(location_zips, getdata))
weather_df_combined <- tbl_df(weather_df_combined)
View(weather_df_combined)
```

*NOTE: The dates used above are TEST DATES to limit calls while experimenting.*

**2) Clean Variables/Column Names/Observations for Joining Dataframes** 

```{r}
weather <- weather_df_combined %>%
  select('date', 'temp', 'cond') %>%
  separate("date", c("date", "time"), sep = " ") %>%
  separate("date", c("year", "month", "day"), sep = "-") %>%
  separate("time", c("hour"), sep = ":") %>% #discarded minutes for join
  rownames_to_column("zipcode")
# adding row names to use for join
weather$zipcode <- substr(weather$zipcode, 1, 5)
# substring row names to use for join
View(weather)
```

```{r}
property <- property %>% 
  filter(ZipName == "Phoenix, AZ" & Month >= "2015-10-01") %>%
  select(Month, zipcode = ZipCode, median_value = "Median Listing Price", 
         total_listed = "Total Listing Count") %>%
  separate("Month", c("year", "month", "day"), sep = "-") %>% 
  select(-"day")
View(property)
```

```{r}
crimes <- crimes %>%
  select(-"OCCURRED TO") %>% 
  separate("OCCURRED ON", c("start_date", "start_time"), sep = "  ") %>%
  separate("start_date", c("month", "day", "year"), sep = "/") %>%
  separate("start_time", c("hour", "minute"), sep = ":") %>% 
  rename(category = "UCR CRIME CATEGORY", block = "100 BLOCK ADDR", 
         zipcode = "ZIP", premise = "PREMISE TYPE")
crimes$zipcode <- as.character(crimes$zipcode)
# change zipcode to chr string for join
View(crimes)
```

**3) Join Dataframes**

```{r}
crimes2 <- left_join(crimes, property)
View(crimes2)
```

*Property data can be obtained through the month prior.*

```{r}
crimes3 <- left_join(crimes2, weather)
View(crimes3)
```

*Weather data after 11/2/15 has not been called.These are test dates, to be replaced with actual dates. I do not know why there are missing values for weather as zipcodes fm crime df were used to specify the call.*

**Coming Soon:**
 1. Might create dummy variables for premise and crime category.
 2. Determine why there are missing weather values. 
 3. Run Rwunderground call for all dates.
 4. Fill missing values.

