---
title: "Statistical Analysis"
author: "Lyttia Cancinos-McManus"
date: "September 2, 2018"
output: html_document
---

0. Load Packages:
```{r}
library(ggplot2)
library(tidyverse)
```
  
0a. Load Data:
```{r}
crimes <- file.choose()
crimes <- read_csv(crimes)

# parsing corrections:
crimes$monthRC <- factor(x = crimes$month, 
                       labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                  "Jul" ,"Aug" ,"Sep" , "Oct", "Nov", "Dec"),
                       ordered = T)
crimes <- crimes %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate_if(is.integer, as.factor)

```

Counts:
_We have a total of 159,319 crimes that have been reported to Phoenix Police Department from Phoenix zip-codes between November 1st, 2015 to June 25th, 2018._

# Exploratory Analysis

_How many crimes occur each month in 2016 and 2017?_

*Counting crime per month: Number of crimes each month in 2016 and 2017.* 
```{r}
crimes2016 <- crimes %>% filter(year == 2016)
crimes2016 %>% count(monthRC)
crimes2017 <- crimes %>% filter(year == 2017)
crimes2017 %>% count(monthRC)

months2016RC <- table(crimes2016$monthRC)
barplot(months2016RC)

months2017RC <- table(crimes2017$monthRC)
barplot(months2017RC)

barplot(months2016RC[order(months2016RC, decreasing = TRUE)])
barplot(months2017RC[order(months2017RC, decreasing = TRUE)])

# Tally by month the dplyr way
group_by(crimes2016, monthRC) %>%
    summarise(count = n()) %>%
    arrange(desc(count))

group_by(crimes2017, monthRC) %>%
    summarise(count = n()) %>%
    arrange(desc(count))
```

*So far I am not seeing seasonal trends. 
Overall Crime does seem to have increased from 2016 to 2017. How much?* 

```{r}
count(crimes2017)- count(crimes2016)
(count(crimes2017)- count(crimes2016))/count(crimes2016)
```

*There were 2,450 more crimes in 2017 than in 2016, a 4.1% increase.* 

_What was the monthly difference in crime from 2016 to 2017?_

```{r}
months2017RC - months2016RC
barplot(months2017RC - months2016RC)
```

*Crimes increased each month in 2017 except for December in which crimes decreased. December went from being the highest crime month in 2016 to #7 in 2017. I wonder why.*

# Descriptive Plots:

-----------------Setup Theme--------------------------------------------------

```{r}

my_theme <- theme_bw() +
    theme(panel.border = element_blank(),
          axis.ticks = element_blank(),
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.major.y = element_line(size = 0.5),
          axis.title = element_text(colour = "grey40"),
          plot.title = element_text(vjust = 2.0))
```


-----------------Total Crimes by Month 2016---------------------

```{r}
ggplot(data = crimes2016, aes(x = monthRC, fill = category)) + 
    my_theme +
    geom_bar(stat = "count", width = 0.75) +
    geom_text(stat = "count", aes(label=..count..)) +
    ggtitle("Crimes reported to Phoenix PD each Month in 2016") +
    xlab("Month") +
    ylab("Crime Count")
```

----------------Total Crimes by Month 2016 (order decr)-----------------------

```{r}
orderedmonths2016 <- dplyr::count(crimes2016, monthRC) %>% 
  arrange(-n) %>% 
  mutate(monthRC = factor(monthRC, monthRC)) 
orderedmonths2016_list <- c("Dec", "Aug", "Oct", "Jan", "Nov", "Apr", "Jun", "Sep", "May", "Jul", "Mar", "Feb")
```

```{r}
ggplot(crimes2016, aes(x = monthRC, fill = category)) + 
  geom_bar(stat = "count", width = 0.75) + 
  scale_x_discrete (limits = orderedmonths2016_list) +
  geom_text(stat = "count", aes(label=..count..), vjust=-0.5) +
  my_theme +
  ggtitle("Crimes reported to Phoenix PD each Month in 2016") +
  xlab("Month (ordered by crime count)") +
  ylab("Crime Count")

#ggsave("crime_count_by_month2016.png")
```

-----------------Total Crimes by Month 2017---------------------

```{r}
ggplot(data = crimes2017, aes(x = monthRC, fill = category)) + 
    my_theme +
    geom_bar(stat = "count", width = 0.75) +
    geom_text(stat = "count", aes(label=..count..)) +
    ggtitle("Crimes reported to Phoenix PD each Month in 2017") +
    xlab("Month") +
    ylab("Crime Count")
```

----------------Total Crimes by Month 2017 (order decr)-----------------------

```{r}
orderedmonths <- dplyr::count(crimes2017, monthRC) %>% 
  arrange(-n) %>% 
  mutate(monthRC = factor(monthRC, monthRC)) 
orderedmonths_list <- c("Oct", "Aug", "Jul", "May", "Jan", "Jun", "Dec", "Mar", "Nov", "Apr", "Sep", "Feb")
```

```{r}
ggplot(crimes2017, aes(x = monthRC, fill = category)) + 
  geom_bar(stat = "count", width = 0.75) + 
  scale_x_discrete (limits = orderedmonths_list) +
  geom_text(stat = "count", aes(label=..count..), vjust=-0.5) +
  my_theme +
  ggtitle("Crimes reported to Phoenix PD each Month in 2017") +
  xlab("Month (ordered by crime count)") +
  ylab("Crime Count")

#ggsave("crime_count_by_month2017.png")
```

# Exploratory Analysis

_How many zip-codes do we have?_

_Counting Zipcodes_

*show zipcodes* 

```{r}
crimes %>% group_by(zipcode) %>% summarise()
```
*Using 43 zipcodes total for analysis out of orig 102 from PPD crime data*

_How many crimes are reported for each zip-code?_

*Counting crimes per zip-code: to return a count of the number of crimes reported in each zip-code*
```{r}
crime_count_by_zip <- crimes %>% 
  count(zipcode)
crime_count_by_zip
```

_Lets see the distribution of the crime categories in each zip-code._ 

# Descriptive Plot:

*Crime count per zipcode, colored by crime type.*

```{r}
g1.5 <- ggplot(crimes, aes(zipcode))
g1.5 + geom_bar(aes(fill=category), width = 0.5) + 
  theme(axis.text.x= element_text(angle = 65, vjust= 0.6)) + 
  geom_text(stat = "Count", aes(label=..count..), vjust=-0.5)+
  ggtitle("Zipcode Frequency Bar Chart") +
  xlab("Zipcode") +
  ylab("Crime Count")
#ggsave("Zipcode Frequency Bar Chart- PHX zips only.png")
```

_Crime count by zip in 2016_

```{r}
g <- ggplot(crimes2016, aes(zipcode, decreasing = TRUE))
g + geom_bar(aes(fill=category), width = 0.5) + 
  theme(axis.text.x= element_text(angle = 65, vjust= 0.6)) + 
  geom_text(stat = "Count", aes(label=..count..), vjust=-0.5)+
  ggtitle("2016 Zipcode Frequency Bar Chart") +
  xlab("Zipcode") +
  ylab("Crime Count")
#ggsave("2016 Zipcode Frequency Bar Chart.png")
```

_Crime count by zip in 2017_

```{r}
g <- ggplot(crimes2017, aes(zipcode, decreasing = TRUE))
g + geom_bar(aes(fill=category), width = 0.5) + 
  theme(axis.text.x= element_text(angle = 65, vjust= 0.6)) +
  geom_text(stat = "count", aes(label=..count..), vjust=-0.5)+
  ggtitle("2017 Zipcode Frequency Bar Chart") +
  xlab("Zipcode") +
  ylab("Crime Count")
#ggsave("2017 Zipcode Frequency Bar Chart.png")
# space the x axis
```

_Which zip-codes are high crime count and which are low?_

#Exploratory Ordered Count

*Zipcodes ordered from highest crime count to lowest.*
```{r}
zipcodes <- table(crimes$zipcode)
ordered_zipcodes <- order(zipcodes, decreasing = TRUE)
ordered_zipcodes
barplot(zipcodes[order(zipcodes, decreasing = TRUE)]) 
# need to make readable
```

_Compare 2016 and 2017_

```{r}
zipcodes2016 <- table(crimes2016$zipcode)
zipcodes2016
ordered_zipcodes2016 <- order(zipcodes2016, decreasing = TRUE)
ordered_zipcodes2016
barplot(zipcodes2016[order(zipcodes2016, decreasing = TRUE)]) 
# need to make readable

zipcodes2017 <- table(crimes2017$zipcode)
zipcodes2017
ordered_zipcodes2017 <- order(zipcodes2017, decreasing = TRUE)
ordered_zipcodes2017
barplot(zipcodes2017[order(zipcodes2017, decreasing = TRUE)]) 
# need to make readable
```

*85015 was the highest crime count zipcode for both years, and crime increased from 2016-2017*

_Which zip codes seem to be getting better, worse?_

```{r}
zipcodes2017-zipcodes2016
barplot(zipcodes2017-zipcodes2016)
Zip_Percentage_change <- (zipcodes2017-zipcodes2016)/zipcodes2016
barplot(Zip_Percentage_change)
# need to make these zipcodes readable
```
*It would be good to note which zipcode drastically increased its crime rate and which zipcode decreased its crime rate.*

_How many crimes occur in the highest crime count zip each month in 2016 and 2017?_

*to return the count of the number of crimes reported each month in the highest crime zip (85015) in 2016*
```{r}
phx15_16 <- crimes %>% 
  filter(zipcode == 85015 & year == 2016) 
  
phx15_16n <- phx15_16 %>% count(monthRC)

phx15_16n

#graph it!
#ERROR
#g <- ggplot(phx15_16, aes(MonthRC, decreasing = TRUE))
#g + geom_bar(aes(fill = category), width = 0.5) + 
#  theme(axis.text.x= element_text(angle = 65, vjust= 0.6)) +
#  geom_text(stat = "count", aes(label=..count..), vjust=-0.5)+
#  ggtitle("Crime in Arizona's worst zipcode in 2016") +
#  xlab("Month") +
#  ylab("Crime Count")
#ggsave("Crime in Arizona's worst zipcode in 2016.png")
```

*to return the count of the number of crimes reported each month in the highest crime zip (85015) in 2017*
```{r}
phx15_17n <- crimes %>% 
  filter(zipcode == 85015 & year == 2017) %>% 
  count(monthRC)

phx15_17n
#graph it!
```

_What crime types are most common?_
*Total number of each category of crime.* 
```{r}
crime_count_by_category <- crimes %>% count(category)
crime_count_by_category
categories <- table(crimes$category)
ordered_categories <- categories[order(categories, decreasing = TRUE)]
barplot(ordered_categories)
```

*Total number of each category of crime in zip code with highest crime count.* 
```{r}
crimes85015 <- crimes %>% filter(zipcode == 85015)
crimes85015 %>% count(category)
categories85015 <- table(crimes85015$category)
barplot(categories85015[order(categories85015, decreasing = TRUE)])
```

```{r}
g <- ggplot(crimes, aes(premise, decreasing = TRUE))
g + geom_bar(aes(fill=category), width = 0.5) + 
  theme(axis.text.x= element_text(angle = 65, vjust= 0.6)) + 
  labs(title= "Premise Frequency Bar Chart")
#ggsave("Premise Frequency Bar Chart.png")

#make table to order for ordered bar plot
premise_table <- table(crimes$premise)
ordered_premise <- order(premise_table, decreasing = TRUE)
barplot(premise_table[ordered_premise])
```

_Does overall crime count increase or decrease as property values increase?_

```{r}
#explore
plot(count(lowincomehousing, median_value), lowincomehousing$median_value)
plot(count(middlehousing, median_value), middlehousing$median_value)
plot(count(upperhousing, median_value), upperhousing$median_value)
```

Comparisons:
_Compare the amounts of crimes in each zip code ordered by property value._
Scatter plots:
```{r}
library(ggplot2)
library(plotly)
options(scipen=999)  # turn-off scientific notation like 1e+48
theme_set(theme_bw())  # pre-set the bw theme.
```

```{r}
gg <- ggplot(upperhousing2, aes(x = as.numeric(median_value), y = n, col = category)) + 
  ylim(0, 3500) + 
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE, size = .4) +
  labs(subtitle="Scatterplot", 
       y="Crime Count", 
       x="Property Value", 
       title="Crime Count by Property Value", 
       caption = "Source: Phoenix Open Data")

plot(gg)
#interactive plotly
#ggplotly(gg)
ggsave("scaled_crime_count_by_property_value_upper.png")
```

```{r}
gg2 <- ggplot(middlehousing2, aes(x = as.numeric(median_value), y = n, color = category)) + 
  geom_point(alpha = 0.4) + 
  geom_smooth(se = FALSE, size = .4) +
  labs(subtitle="Scatterplot", 
       y="Crime Count", 
       x="Property Value", 
       title="Crime Count by Property Value", 
       caption = "Source: Phoenix Open Data")

plot(gg2)
#interactive plotly
#ggplotly(gg2)
ggsave("crime_count_by_property_value_middle.png")
```


```{r}
gg3 <- ggplot(lowincomehousing2, aes(x = as.numeric(median_value), y = n, color = category)) + 
  ylim(0, 3500) +
  geom_point(alpha = 0.4) + 
  geom_smooth(se = FALSE, size = .4) +
  labs(subtitle="Scatterplot", 
       y="Crime Count", 
       x="Property Value", 
       title="Crime Count by Property Value", 
       caption = "Source: Phoenix Open Data")

plot(gg3)
#interactive plotly
#ggplotly(gg3)
ggsave("crime_count_by_property_value_low.png")
```

Time-series:

```{r}
#install.packages("plotly")
#library(plyr)
#library(dplyr)
library(plotly)

crimes2016_2017 <- filter(crimes, year %in% c("2016","2017"))

total_crime_2years <-
  crimes2016_2017 %>% 
  mutate(year = as.integer(as.character(year))) %>%
  group_by(crime.type, year) %>% 
  dplyr::summarize(reported_incidents = n())

trend_plot <- 
  ggplot(data = total_crime_2years,
       aes(x = year, y = reported_incidents, fill = crime.type)) +
  geom_area() +
  ggtitle("Violent vs. Non-Violent Crimes in 2016-2017") +
  scale_y_continuous(name = "Reported Incidents", labels = scales::comma) +
  my_theme

ggplotly(trend_plot)

totalcat_crime_2years <-
  crimes2016_2017 %>% 
  mutate(year = as.integer(as.character(year))) %>%
  group_by(category, year) %>% 
  dplyr::summarize(reported_incidents = n())

trend_plot2 <- 
  ggplot(data = totalcat_crime_2years,
       aes(x = year, y = reported_incidents, fill = category)) +
  geom_area() +
  theme(legend.position='bottom')
  ggtitle("Crime Categories in 2016-2017") +
  scale_y_continuous(name = "Reported Incidents", labels = scales::comma) +
  my_theme

ggplotly(trend_plot2)

```


# Insights/ correlations:

* Trends (high, low, increase, decrease, anomalies):

- Zip code 85015 has the highest crime count total at almost 9000 crimes (8965).

- Lowest crime count total: 

- Total crime count seems to be between 5,000 - 6,000 each month in 2016 & 2017.

- Overall crime has increased from one year to the next. 

- Most zip codes reported more crime in 2017, but several reported less: 

- Highest decrease in crime from 2016 - 2017:

- Highest increase in crime from 2016 - 2017:

- December decreased in crime count drastically from 2016 to 2017. From 1st to 7th place. 

- Lowest count month: 

- Crime count peaks at lower middle housing, and as property value increased, crime count decreased.

- Most common place for crime to occur in Phoenix are single family homes, with burglary being the most common crime. 

* Is there a hypothesis you would like to investigate further? 

- There is likely a daily or hourly trend.... lets plot this next.

- 

* What other questions do they lead you to ask?

-Do people in 85015 simply report more crimes than other phx zipcodes? Or are there actually more crimes?

# Limitations: 

- The data accounts only for crimes reported to Phoenix PD. There may be many other crimes reported to surrounding PDs.
- Not enough variables. 
- Not enough years of data. 
- No data on area demographics, criminals, victims, crime status, or repeat crimes/criminals/victims. 
