---
title: "Rwunderground Call/ Cleaning Script"
output: html_notebook
---
# Rwunderground Call/ Cleaning Script

*I am using an R interface to weather underground's API. In order to use this library you must obtain an API key. In the past, a free key was available. The free key limits requests to 10 per minute. If you are grabbing weather for a large date range using history_range then by default limit = 10 will limit the calls to a maximum of 10 per minute. This package has functions that follow the online api.*

**0- Install & Load packages**

```{r}
#install.packages("tidyverse", dependencies = TRUE)
#install.packages("devtools")
#install.packages("stringi", dependencies = TRUE)
#devtools::install_github("ALShum/rwunderground", force = TRUE, dependencies = TRUE)
library(tidyverse)
library(devtools)
library(dplyr)
library(rwunderground)
# Set API key
rwunderground::set_api_key("337b40a1234484ca")
WUNDERGROUNDID = '337b40a1234484ca'
```

**1a- Create function to Call rwundergound API Package and produce PHX weather history df**

```{r}
#change zipcode assignment
zip_code = as.character(unique(crimes$ZIP))
View(zip_code)

#change zipcodes
location_zips = sapply(zip_code, set_location)
View(location_zips)
```


```{r}
getdata <- function(location_zip, date_start = "20151101", 
                    date_end = "20151102"){
                    df = history_range(location_zip, date_start, date_end, 
                                       limit = 10) 
                    return (df)
                    }
getdata2 <- function(location_zip, date_start = "20151103", 
                    date_end = "20151104"){
                    df  = history_range(location_zip, date_start, date_end, 
                                        limit = 10)
                    return (df)
                    }
getdata3 <- function(location_zip, date_start = "20151105", 
                    date_end = "20151107"){
                    df  = history_range(location_zip, date_start, date_end, 
                                        limit = 10)
                    return (df)
                    }
getdata4 <- function(location_zip, date_start = "20151108", 
                    date_end = "20151110"){
                    df  = history_range(location_zip, date_start, date_end, 
                                        limit = 10)
                    return (df)
                    }
getdata5 <- function(location_zip, date_start = "20151111", 
                    date_end = "20151115"){
                    df  = history_range(location_zip, date_start, date_end, 
                                        limit = 10)
                    return (df)
                    }
getdata6 <- function(location_zip, date_start = "20151116", 
                    date_end = "20151130"){
                    df  = history_range(location_zip, date_start, date_end, 
                                        limit = 10)
                    return (df)
                    }
getdata7 <- function(location_zip, date_start = "20151201", 
                    date_end = "20151215"){
                    df  = history_range(location_zip, date_start, date_end, 
                                        limit = 5) 
                    return (df)
                    }
getdata7.5 <- function(location_zip, date_start = "20151216",
                       date_end = "20151231"){
  df  = history_range(location_zip, date_start, date_end, 
                      limit = 6)
  return (df)
  }
getdata8 <- function(location_zip, date_start = "20160101", 
                       date_end = "20160115"){
  df  = history_range(location_zip, date_start, date_end, 
                      limit = 5)
  return (df)
}
getdata8.5 <- function(location_zip, date_start = "20160116", 
                     date_end = "20160131"){
  df  = history_range(location_zip, date_start, date_end, 
                      limit = 5)
  return (df)
}
getdata9 <- function(location_zip, date_start = "20160201", 
                     date_end = "20160215"){
  df  = history_range(location_zip, date_start, date_end, 
                      limit = 5)
  return (df)
}
getdata9.5 <- function(location_zip, date_start = "20160216", 
                     date_end = "20160229"){
  df  = history_range(location_zip, date_start, date_end, 
                      limit = 5)
  return (df)
}
getdata10 <- function(location_zip, date_start = "20160301", 
                     date_end = "20160315"){
  df  = history_range(location_zip, date_start, date_end, 
                      limit = 5)
  return (df)
}
getdata10.5 <- function(location_zip, date_start = "20160316", 
                      date_end = "20160331"){
  df  = history_range(location_zip, date_start, date_end, 
                      limit = 5)
  return (df)
}
getdata11 <- function(location_zip, date_start = "20160401", 
                      date_end = "20160415"){
  df  = history_range(location_zip, date_start, date_end, 
                      limit = 5)
  return (df)
}
getdata11.5 <- function(location_zip, date_start = "20160416", 
                      date_end = "20160430"){
  df  = history_range(location_zip, date_start, date_end, 
                      limit = 5)
  return (df)
}

getdata12 <- function(location_zip, date_start = "20160501", 
                      date_end = "20160515"){
  df  = history_range(location_zip, date_start, date_end, 
                      limit = 5)
  return (df)
}

getdata12.5 <- function(location_zip, date_start = "20160516", 
                      date_end = "20160530"){
  df  = history_range(location_zip, date_start, date_end, 
                      limit = 5)
  return (df)
}
#---------------------->>>>>>>>>> START HERE <<<<<<<<<<----------------------------
```


```{r}
weather_df_combined = do.call(rbind, lapply(location_zips, getdata))
weather_df_combined2 = do.call(rbind, lapply(location_zips, getdata2))
weather_df_combined3 = do.call(rbind, lapply(location_zips, getdata3))
weather_df_combined4 = do.call(rbind, lapply(location_zips, getdata4))
weather_df_combined5 = do.call(rbind, lapply(location_zips, getdata5))
weather_df_combined6 = do.call(rbind, lapply(location_zips, getdata6))
#getdata6 stopped at index 27(zipcode: 85016) on 11/24
weather_df_combined6.5 = do.call(rbind, lapply(location_zips[27:102], getdata6))
weather_df_combined7 = do.call(rbind, lapply(location_zips, getdata7))
weather_df_combined7.5 = do.call(rbind, lapply(location_zips, getdata7.5))
#getdata7.5 stopped at index 38 (85021.121) on 12/20
weather_df_combined7.75 = do.call(rbind, lapply(location_zips[39:102], getdata7.5))
weather_df_combined8 = do.call(rbind, lapply(location_zips, getdata8))
#----------------->>>>>>>>>> Check for missing values <<<<<<<<<<-----------------------
weather_df_combined8.5 = do.call(rbind, lapply(location_zips, getdata8.5))
# stopped at (85250)
weather_df_combined9 = do.call(rbind, lapply(location_zips, getdata9))
#stopped at 
weather_df_combined9.5 = do.call(rbind, lapply(location_zips, getdata9.5))
#stopped at (85250)
weather_df_combined10 = do.call(rbind, lapply(location_zips, getdata10))
weather_df_combined10.5 = do.call(rbind, lapply(location_zips, getdata10.5))
weather_df_combined11 = do.call(rbind, lapply(location_zips, getdata11))
weather_df_combined11.5 = do.call(rbind, lapply(location_zips, getdata11.5))
weather_df_combined12 = do.call(rbind, lapply(location_zips, getdata12))
# stopped at: 20160504/q/85305
#---------------------->>>>>>>>>> START HERE<<<<<<<<<<----------------------------
rm(list=ls(pattern = "^getdata"))
```

```{r}
w1 <- tbl_df(weather_df_combined)
w2 <- tbl_df(weather_df_combined2)
w3 <- tbl_df(weather_df_combined3)
w4 <- tbl_df(weather_df_combined4)
w5 <- tbl_df(weather_df_combined5)
w6 <- tbl_df(weather_df_combined6)
w6.5 <- tbl_df(weather_df_combined6.5)
w7 <- tbl_df(weather_df_combined7)
w7.5 <- tbl_df(weather_df_combined7.5)
w7.75 <- tbl_df(weather_df_combined7.75)
w8 <- tbl_df(weather_df_combined8)
w8.5 <- tbl_df(weather_df_combined8.5)
w9 <- tbl_df(weather_df_combined9)
w9.5 <- tbl_df(weather_df_combined9.5)
w10 <- tbl_df(weather_df_combined10)
w10.5 <- tbl_df(weather_df_combined10.5)
w11 <- tbl_df(weather_df_combined11)
w11.5 <- tbl_df(weather_df_combined11.5)
w12 <- tbl_df(weather_df_combined12)
rm(list = ls(pattern = "^weather_df_combined"))
gc()
```

**1b- Join Weather dataframes**

```{r}
weather_2015dfs <- rbind(w1, w2, w3,w4, w5, w6, w6.5, w7, w7.5, w7.75)
View(weather_2015dfs)
weather_2016dfs <- rbind(w8, w8.5, w9, w9.5, w10, w10.5, w11, w11.5, w12)
```

**2- Clean Variables/Column Names/Observations**

```{r}
weather2015 <- weather_2015dfs %>%
  select('date', 'temp', 'cond') %>%
  separate("date", c("date", "time"), sep = " ") %>%
  separate("date", c("year", "month", "day"), sep = "-") %>%
  separate("time", c("hour"), sep = ":") %>% #discarded minutes for join
  rownames_to_column("zipcode")
# adding row names to use for join
weather2015$zipcode <- substr(weather2015$zipcode, 1, 5)
# substring row names to use for join

View(weather2015)

weather2016 <- weather_2016dfs %>%
  select('date', 'temp', 'cond') %>%
  separate("date", c("date", "time"), sep = " ") %>%
  separate("date", c("year", "month", "day"), sep = "-") %>%
  separate("time", c("hour"), sep = ":") %>% #discarded minutes for join
  rownames_to_column("zipcode")
# adding row names to use for join
weather2016$zipcode <- substr(weather2016$zipcode, 1, 5)
# substring row names to use for join

View(weather2016)

rm(list = ls(pattern = "^..."))
```
